### 2018-06-04 Writing/draft_highlighted.tex
new version draft

@@

\newpage
\begin{abstract}
Phylogenetic analysis algorithms require the assumption of character independence - a condition generally acknowledged to be violated by morphological data.
Correlation between characters can originate from intra-organismal features, shared phylogenetic history or forced by particular character-state coding schemes. 
Although the two first sources can be investigated by biologists \textit{a posteriori} and the third one can be avoided \textit{a priori} with good practices, phylogenetic software do not distinguish between any of them.

@@

    \item \textbf{Evolutionary dependence:} this is the result of sets of characters co-evolving due to selection, likely related to functional links between two traits that help serve an overall lifestyle trait. 
    Unlike the case of intra-organismal dependence, there need not be an intrinsic constraint that causes these traits to covary.
    For example, in vertebrates, axial elongation can be correlated to limb reduction with snake-like bodies evolving multiple times in numerous tetrapod lineages.
    This is thought to correspond to adaptations for fossoriality or aquatic lifestyles.
    Such covariances are generally studied in the context of a given phylogeny, often one derived from molecular data with the morphological traits of interest mapped on it.
    Many methods have been developed to study these correlations, especially since they can provide us with a lot of information on how specific groups acquired specific characteristics \citep{Lande1983,Maddison1990,Pagel1994,Pagel2006,Grabowski2016}.
    However, these methods do not give us a means to objectively control correlations that might adversely affect phylogenetic inference.

    \item \textbf{Coding dependence:} this is the results of researcher methodology for defining or/and coding discrete morphological characters \citep{Brazeau2011,simoes2017giant}.
    Coding dependence manifests itself in several ways, particularly in coding redundant information.
    For instance, coding for the same absence in different characters creates state transformations associated with the loss or gain of a particular character.
    This occurs when a number of multistate characters include two variable feature states (e.g. large, small; red, blue etc.) in conjunction with absence.
    It is worth noting, however, that these correlations could also be due to the nature of the available data, especially in palaeontology.
    For example, when only one fragmentary molar is available to describe a specimen, researchers have to ``extract'' as much phylogenetic information from the available data as possible, potentially inducing correlations.
    This coding dependency is linked to hierarchical dependency between characters \citep{wilkinson1995coping,BrazeauNA}.
    Finally this can also be due to a bias in the amount of characters available.
    For example, in skulls, because of their complexity, there is a high likelihood of inducing correlation (by effectively reducing structural complexity to discrete characters).
\end{itemize}

@@

\noindent Where $n$ is the number of taxa with comparable characters $x$, $y$ and $x_i$, $y_i$ are each character's state for the $i^{th}$ taxon.
$CD$ is a continuous distance metric bounded between $0$ and $1$ (see the mathematical demonstration in the supplementary material 1).
Since we are considering differences as being only Fitch-like (non-additive) and unweighted, we calculated the difference between character states in a qualitative way.
Two same character states tokens have a difference of zero and two different ones have a difference of one (e.g. $0 - 0 = 0$ or $1 - 8 = 1$).
Additionally, we only consider differences for taxa with shared information \citep[i.e. a Gower distance;][]{GowerDist}.

@@

\subsection{Simulating discrete morphological matrices}
To simulate the matrices we applied a protocol very similar to \cite{Guillerme2016146}.
First, we generate random birth-death trees with the birth ($\lambda$) and death ($\mu$) parameters sampled from a uniform $(0,1)$ distribution maintaining $\lambda$ $>$ $\mu$ using the \texttt{diversitree} \texttt{R} package \citep[v0.9-8;][]{fitzjohndiversitree2012} and saving the tree after reaching either 25, 75 or 150 taxa.
For each tree, we arbitrarily set the outgroup to be the first taxon (alphabetically) thus effectively rooting the trees on this taxon.
These trees are hereafter called the ``true'' trees (see distinction below).
We then simulated discrete morphological characters on the topology of these trees using the either of the two following models:
\begin{itemize}
    \item The morphological HKY-binary model \citep{OReilly20160081} which is an HKY model \citep{HKY85} with a random states frequency (sampled from a Dirichlet distribution $Dir(1,1,1,1)$) and using a transition/transvertion rate of $2$ \citep{douadycomparison2003} but where the purines (A,G) were changed into state $0$ and the pyrimidines (C,T) in state $1$.
    This model has the advantage of not favouring Bayesian inference \citep[since it doesn't use an M$k$ model;][; see discussion]{OReilly20160081} but the downside of it is it can only generate binary state characters \citep[or 4 states;][]{puttick2017uncertain}.
    \item To generate more than binary states characters, we used the M$k$ model \citep{lewisa2001}.
    We draw the number of character states with a probability of $0.85$ for binary characters and $0.15$ for three state characters \citep{Guillerme2016146,ZouConvergence}.
    This model assumes a equal transition rate between character states which might seem overly simplistic, excluding other observed transition patterns \citep[e.g. Dollo characters;][]{Dollo,wright2015came}.
    Recently however, \cite{Wright01072016} have shown that an equal rate transition is still the most present in empirical data.
\end{itemize}

\noindent For each character, both models (morphological HKY-binary or M$k$) where chosen randomly and run with an overall evolutionary rate drawn from a gamma distribution ($\beta$ = $100$ and $\alpha$ = $5$).
These low evolutionary rate values allowed reduction in the number of homoplasic character changes, thus reinforcing the phylogenetic information in the matrices.
We re-simulated every invariant characters to obtain a matrix with no invariant characters in order to better approximate real morphological data matrices.
To ensure that our simulations were reflecting realistic observed parameters, we only selected matrices with Consistency Indices (CI) superior to $0.26$ \citep{OReilly20160081}.

@@

The resulting full simulation was 3.5TB big so is not shared here (though the parameters are).
However, the resulting consensus trees on which the topological differences are calculated are available at \url{https://figshare.com/s/7a8fde8eaa39a3d3cf56}.

>>>>

\usepackage{color,soul}

@@

\newpage
\begin{abstract}
Phylogenetic analysis algorithms require the assumption of character independence - a condition generally acknowledged to be violated by morphological data.
Correlation between characters can originate from intra-organismal features, shared phylogenetic history or forced by particular character-state coding schemes. 
Although the two first sources can be investigated by biologists \textit{a posteriori} and the third one can be avoided \textit{a priori} with \hl{careful character coding}, phylogenetic software do not distinguish between any of them.

@@

    \item \textbf{Evolutionary dependence:} this is the result of sets of characters co-evolving due to selection, likely related to functional links between two traits that help serve an overall lifestyle trait \citep{ClarkeMiddleton2008}. 
    Unlike the case of intra-organismal dependence, there need not be an intrinsic constraint that causes these traits to covary.
    For example, in vertebrates, axial elongation can be correlated to limb reduction with snake-like bodies evolving multiple times in numerous tetrapod lineages.
    This is thought to correspond to adaptations for fossoriality or aquatic lifestyles.
    Such covariances are generally studied in the context of a given phylogeny, often one derived from molecular data with the morphological traits of interest mapped on it.
    Many methods have been developed to study these correlations, especially since they can provide us with a lot of information on how specific groups acquired specific characteristics \citep{Lande1983,Maddison1990,Pagel1994,Pagel2006,Grabowski2016}.
    However, these methods do not give us a means to objectively control correlations that might adversely affect phylogenetic inference.

    \item \textbf{Coding dependence:} this is the results of researcher methodology for defining or/and coding discrete morphological characters \citep{Brazeau2011,simoes2017giant}.
    Coding dependence manifests itself in several ways, particularly in coding redundant information.
    For instance, coding for the same absence in different characters creates state transformations associated with the loss or gain of a particular character.
    This occurs when a number of multistate characters include two variable feature states (e.g. large, small; red, blue etc.) in conjunction with absence.
    \hl{For example, two characters ``tail colour'' and ``tail length'' could be coded two times as an absence for a taxon with no tail.}
    It is worth noting, however, that these correlations could also be due to the nature of the available data, especially in palaeontology.
    For example, when only one fragmentary molar is available to describe a specimen, researchers have to ``extract'' as much phylogenetic information from the available data as possible, potentially inducing correlations.
    This coding dependency is linked to hierarchical dependency between characters \citep{wilkinson1995coping,BrazeauNA}.
    Finally this can also be due to a bias in the amount of characters available.
    For example, in skulls, because of their complexity, there is a high likelihood of inducing correlation (by effectively reducing structural complexity to discrete characters).
\end{itemize}

@@

\noindent Where $n$ is the number of taxa with comparable characters $x$, $y$ and $x_i$, $y_i$ are each character's state for the $i^{th}$ taxon.
$CD$ is a continuous distance metric bounded between $0$ and $1$ (see the mathematical demonstration in the supplementary material 1).
Since we are considering differences as being only Fitch-like (non-additive) and unweighted, we calculated the difference between character states in a qualitative way.
Two same character states tokens have a difference of zero and two different ones have a difference of one (e.g. $0 - 0 = 0$ or $1 - 8 = 1$).
Additionally, we only consider differences for taxa with shared information \citep[i.e. a Gower distance;][]{GowerDist}.
\hl{For example, for two characters X and Y for four taxa, if character \texttt{X = \{0,0,?,1\}} for the first, second, third and fourth taxa respectively and character \texttt{Y = \{0,0,1,1\}}, only the first, second and fourth taxa are compared for this character.}

@@

\subsection{Simulating discrete morphological matrices}
To simulate the matrices we applied a protocol very similar to \cite{Guillerme2016146}.
First, we generate random birth-death trees with the birth ($\lambda$) and death ($\mu$) parameters sampled from a uniform $(0,1)$ distribution maintaining $\lambda$ $>$ $\mu$ using the \texttt{diversitree} \texttt{R} package \citep[v0.9-8;][]{fitzjohndiversitree2012} and saving the tree after reaching either 25, 75 or 150 taxa.
For each tree, we arbitrarily set the outgroup to be the first taxon (alphabetically) thus effectively rooting the trees on this taxon.
These trees are hereafter called the ``true'' trees (see distinction below).
We then simulated discrete morphological characters on the topology of these trees using the either of the two following models:
\begin{itemize}
    \item The \hl{``morphological HKY-binary'' model} \citep{OReilly20160081} which is an HKY model \citep{HKY85} with a random states frequency (sampled from a Dirichlet distribution $Dir(1,1,1,1)$) and using a transition/transvertion rate of $2$ \citep{douadycomparison2003} but where the purines (A,G) were changed into state $0$ and the pyrimidines (C,T) in state $1$.
    This model has the advantage of not favouring Bayesian inference \citep[since it doesn't use an M$k$ model;][; see discussion]{OReilly20160081} but the downside of it is it can only generate binary state characters \citep[or 4 states;][]{puttick2017uncertain}.
    \item To generate more than binary states characters, we used the M$k$ model \citep{lewisa2001}.
    We draw the number of character states with a probability of $0.85$ for binary characters and $0.15$ for three state characters \citep{Guillerme2016146,ZouConvergence}.
    This model assumes a equal transition rate between character states which might seem overly simplistic, excluding other observed transition patterns \citep[e.g. Dollo characters;][]{Dollo,wright2015came}.
    However, \cite{Wright01072016} \hl{have shown that equal transition rate is a reasonable assumption for many empirical datasets.}
\end{itemize}

\noindent For each character, both models (``morphological HKY-binary'' or M$k$) where chosen randomly and run with an overall evolutionary rate drawn from a gamma distribution ($\beta$ = $100$ and $\alpha$ = $5$).
These low evolutionary rate values allowed reduction in the number of homoplasic character changes, thus reinforcing the phylogenetic information in the matrices.
We re-simulated every invariant characters to obtain a matrix with no invariant characters in order to better approximate real morphological data matrices.
To ensure that our simulations were reflecting realistic observed parameters, we only selected matrices with Consistency Indices (CI) superior to $0.26$ \citep{OReilly20160081}.
\hl{Note that both models (``morphological HKY-binary' and Mk) are both based on a Markovian model and only differ in the number of states generated and transformed.
However, the ``morphological HKY model'' is not the one \textit{directly}estimated in the phylogenetic software used in this protocol (MrBayes, see below).}

@@

The resulting full simulation was 3.5TB \hl{in size} so is not shared here (though the parameters are).
However, the resulting consensus trees on which the topological differences are calculated are available at \url{https://figshare.com/s/7a8fde8eaa39a3d3cf56}.
